A 38 Million Year Old Mini-Neptune in the Kepler Field

### 2021-09-15 paper/ms.tex
update ms
AU Mic c exists

Kepler\,1627Ab provides a new extremum in the ages of the Kepler
planets, and opens multiple avenues for further study.  Further
observations of Doppler transits could yield a more precise stellar
obliquity, and could confirm or refute the expected prograde orbital
geometry.  Similarly, transit spectroscopy aimed at detecting outflows
of hydrogen or helium could yield insight into the evolutionary state
of the atmosphere \citep[{\it
e.g.},][]{spake_helium_2018,vissapragada_2020}.  Determining the
precise amount of high-energy irradiation incident on the planet would
also help in clarifying the expected outflow rate \citep[{\it
e.g.},][]{poppenhaeger_2021}.  Finally, a challenging but informative
quantity to measure would be the planet's mass.  This might be
possible either through detailed timing analyses of the Kepler data,
or through a high-cadence multi-color radial velocity campaign.
Measured at sufficient precision, the mass could yield constraints on
both the planet's composition and its initial entropy
\citep{owen_constraining_2020}.

Separately, the Kepler data may yet contain further information.
Although our transit timing study did not yield the detection of any
clearly significant planet-planet TTVs, Appendix~\ref{app:flare}
highlights an odd flare timing coincidence that may be suggestive of
star-planet interactions.  We mention it here to encourage its
exploration by investigators more versed in the topic than ourselves.

In the context of the transiting planet population, \pn\ is among the
youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
with precise ages below 100\,Myr include K2-33b, DS~Tuc~Ab,
HIP~67522b, TOI~837b, AU~Mic~b, and the four-planet V1298~Tau system
\citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020}.
\pn\ is currently the smallest of these systems ($3.34\pm
0.18\,R_\oplus$), which may hint at selection effects in the young
planet sample \citep{zhou_2021_tois}.  Assuming that these planets
have masses between $\approx$5$\,M_\oplus$ and 10$\,M_\oplus$, their
large sizes and presumed low densities would be in accord with the
expectation that mini-Neptunes start their lives with large primordial
atmospheres that are shed over the first gigayear
\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.  Our
suggestion of an aligned orbit differs from observed misalignments in
number of close-in
mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
Unlike these older planets, \pn\ seems to have properties consistent
with a quiescent disk-driven migration history.  The alternative
scenario of in-situ formation is disfavored on theoretical grounds:
the required disk mass would likely drive migration, and so lacks
internal consistency
\citep{inamdar_formation_2015,ogihara_reassessment_2015}.

Ultimately, the main advance of this work is a precise measurement of
the age of \pn.   This age measurement was enabled by identifying the
connection of the star to the \cn\ using the Gaia data.
Table~\ref{tab:v06} should enable similar cross-matches to already
known planets, as well as forthcoming planets from the TESS mission
\citep{ricker_transiting_2015,guerrero_tess_2021}.  An alternative
path toward age-dating more planets could be to identify kinematic
associations around known exoplanet host stars using positions and
tangential velocities from Gaia, and to then verify these associations
with ancillary rotation periods and spectroscopy \citep[{\it
e.g.},][]{tofflemire_tess_2021}.

@@

The authors are grateful to T{.}~David, J{.}~Winn, and A{.}~Howard for
illuminating discussions and guidance, and to R{.}~Kerr for kindly
providing us with the \citet{Kerr2021} membership list prior to its
publication.
The authors are also grateful to K{.}~Collins for
helping resolve the scheduling conflict that would have otherwise
prevented the MuSCAT3 observations.
% is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
% E{.}~Newton, M{.}~Kounkel, A{.}~Kraus, L{.}~Hillenbrand, and
% K{.}~Hawkins for the discussions on young stars, rotation, and lithium
% that encouraged this analysis.
%
L.G.B{.} acknowledges support from a Charlotte Elizabeth Procter
Fellowship from Princeton University, as well as from the TESS GI Program, programs
G011103 and G022117, through NASA grants 80NSSC19K0386 and
80NSSC19K1728.
%
%
Keck/NIRC2 imaging was acquired by programs 2015A/N301N2L
(PI: A.~Kraus) and 2019A/N069 (PI: E.~Petigura).
%
In addition, this paper is based in part on observations made with the
MuSCAT3 instrument, developed by the Astrobiology Center and under
financial support by JSPS KAKENHI (JP18H05439) and JST PRESTO
(JPMJPR1775), at Faulkes Telescope North on Maui, HI, operated by the
Las Cumbres Observatory.
%
This work is partly supported by JSPS KAKENHI Grant Numbers 22000005,
JP15H02063, JP17H04574, JP18H05439, JP18H05442, JST PRESTO Grant
Number JPMJPR1775, the Astrobiology Center of National Institutes of
Natural Sciences (NINS) (Grant Number AB031010).
%
% ACKNOWLEDGE PFS / CAMPANAS.
%
This paper also includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
We thank the TESS Architects (G.~Ricker, R.~Vanderspek, D.~Latham,
S.~Seager, J.~Jenkins) and the many TESS team members for their
efforts to make the mission a continued success.
%
%
%This study was based in part on observations at Cerro Tololo
%Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
%2020A-0146; 2020B-0029 PI: Bouma), which is managed by the
%Association of Universities for Research in Astronomy (AURA) under a
%cooperative agreement with the National Science Foundation.
%
%
Finally, this research has made use of the Keck Observatory Archive (KOA),
which is operated by the W. M. Keck Observatory and the NASA Exoplanet
Science Institute (NExScI), under contract with the National
Aeronautics and Space Administration.  We also thank the Keck
Observatory staff for their support of HIRES and remote observing.  We
recognize the importance that the summit of Maunakea has within the
indigenous Hawaiian community, and are deeply grateful to have the
opportunity to conduct observations from this mountain.
%
% The Digitized Sky Survey was produced at the Space Telescope Science
% Institute under U.S. Government grant NAG W-2166.
% Figure~\ref{fig:scene} is based on photographic data obtained using
% the Oschin Schmidt Telescope on Palomar Mountain.
%

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.9\textwidth]{f13.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Statistics of inter-flare arrival times}.  
    24 flares were recorded with amplitudes exceeding 0.5\% over the
    97.7 days of short cadence observations.  The histogram of the
    time intervals between every possible pair of flares is shown in
    black.  Some plausibly important timescales for star-planet
    interactions, namely the planetary orbital period and synodic
    period (the orbital period as seen from the rotating stellar frame) are
    shown along with their linear combinations.  Monte Carlo
    draws from a Poisson distribution are shown
    with the gray bands.  While peaks in the observed distribution do
    coincide with some of the ``special periods'', the
    statistical evidence for a non-Poissonian process driving the
    flares does not reach the $5$-$\sigma$ threshold.
		\label{fig:flarestats}
	}
\end{figure*}

>>>>

Kepler\,1627Ab provides a new extremum in the ages of the Kepler
planets, and opens multiple avenues for further study.  Further
observations of Doppler transits could yield a more precise stellar
obliquity, and could confirm or refute the expected prograde orbital
geometry.  Similarly, transit spectroscopy aimed at detecting outflows
of hydrogen or helium could yield insight into the evolutionary state
of the atmosphere \citep{spake_helium_2018,vissapragada_2020}.  Determining the
precise amount of high-energy irradiation incident on the planet would
also help in clarifying the expected outflow rate \citep[{\it
e.g.},][]{poppenhaeger_2021}.  Finally, a challenging but informative
quantity to measure would be the planet's mass.  This might be
possible through timing analyses of the Kepler data,
or through a high-cadence multi-color radial velocity campaign.
Measured at sufficient precision, the mass could yield constraints on
both the planet's composition and its initial entropy
\citep{owen_constraining_2020}.

Separately, the Kepler data may yet contain further information.
Although our transit timing study did not yield the detection of any
clearly significant planet-planet TTVs, Appendix~\ref{app:flare}
highlights an odd flare timing coincidence that may be suggestive of
star-planet interactions.  We encourage its
exploration by investigators more versed in the topic than ourselves.

In the context of the transiting planet population, \pn\ is among the
youngest known (Figure~\ref{fig:rp_period_age}).  Comparable systems
with precise ages below 100\,Myr include K2-33, DS~Tuc~A, HIP~67522,
TOI~837, the two-planet AU~Mic system, and the four-planet V1298~Tau
system
\citep{Mann_K2_33b_2016,David_et_al_2017,newton_tess_2019,david_four_2019,bouma_cluster_2020,plavchan_planet_2020,martioli_aumicbc_2021}.
\pn\ is one of the smallest planets in this sample ($3.34\pm
0.18\,R_\oplus$), which could be linked to the selection effects
imposed by the spot-induced photometric variability at very young ages
\citep[{\it e.g.},][]{zhou_2021_tois}.  Assuming that these young
planets have masses between $\approx$5$\,M_\oplus$ and 10$\,M_\oplus$,
their large sizes and presumed low densities would be in accord with
the expectation that mini-Neptunes start their lives with large
primordial atmospheres that are shed over the first gigayear
\citep{Owen_Wu_2013,Fulton_et_al_2017,ginzburg_corepowered_2018}.  Our
suggestion of an aligned orbit is consistent with a quiescent
disk-driven migration history, and differs from misalignments that
have been observed for some older
mini-Neptunes~\citep{sanchis-ojeda_starspots_2011,albrecht_obliquities_2012,dalal_2019_hd3167,rubenzahl_tess-keck_2021}.
A more detailed spectroscopic analysis of the transit would however be
worthwhile to verify whether the TTV-slope correlation is yielding the
correct orbital geometry.
%The alternative
%scenario of in-situ formation is disfavored on theoretical grounds:
%the required disk mass would likely drive migration, and so lacks
%internal consistency
%\citep{inamdar_formation_2015,ogihara_reassessment_2015}.

Ultimately, the main advance of this work is a precise measurement of
the age of \pn.   This age measurement was enabled by identifying the
connection of the star to the \cn\ using the Gaia data.
Table~\ref{tab:v06} should enable similar cross-matches to already
known planets, as well as forthcoming planets from the TESS mission
\citep{ricker_transiting_2015,guerrero_tess_2021}.  An alternative
approach for age-dating more planets could be to identify kinematic
associations around known exoplanet host stars using positions and
tangential velocities from Gaia, and to then verify these associations
with ancillary rotation periods and spectroscopy \citep[{\it
e.g.},][]{tofflemire_tess_2021}.

@@

The authors are grateful to T{.}~David, J{.}~Winn, and A{.}~Howard for
illuminating discussions and guidance, and to R{.}~Kerr for kindly
providing us with the \citet{Kerr2021} membership list prior to its
publication.
The authors are also grateful to K{.}~Collins for
helping resolve the scheduling conflict that would have otherwise
prevented the MuSCAT3 observations.
% is grateful to G.~Zhou, B{.}~Tofflemire, A{.}~McWilliam,
% E{.}~Newton, M{.}~Kounkel, A{.}~Kraus, L{.}~Hillenbrand, and
% K{.}~Hawkins for the discussions on young stars, rotation, and lithium
% that encouraged this analysis.
%
L.G.B{.} acknowledges support from a Charlotte Elizabeth Procter
Fellowship from Princeton University, as well as from the TESS GI Program, programs
G011103 and G022117, through NASA grants 80NSSC19K0386 and
80NSSC19K1728.
%
%
Keck/NIRC2 imaging was acquired by program 2015A/N301N2L
(PI: A.~Kraus). % and 2019A/N069 (PI: E.~Petigura).
%
In addition, this paper is based in part on observations made with the
MuSCAT3 instrument, developed by the Astrobiology Center and under
financial support by JSPS KAKENHI (JP18H05439) and JST PRESTO
(JPMJPR1775), at Faulkes Telescope North on Maui, HI, operated by the
Las Cumbres Observatory.
%
This work is partly supported by JSPS KAKENHI Grant Numbers 22000005,
JP15H02063, JP17H04574, JP18H05439, JP18H05442, JST PRESTO Grant
Number JPMJPR1775, the Astrobiology Center of National Institutes of
Natural Sciences (NINS) (Grant Number AB031010).
%
% ACKNOWLEDGE PFS / CAMPANAS.
%
This paper also includes data collected by the TESS mission, which are
publicly available from the Mikulski Archive for Space Telescopes
(MAST).
%
Funding for the TESS mission is provided by NASA's Science Mission
directorate.
%
We thank the TESS Architects (G.~Ricker, R.~Vanderspek, D.~Latham,
S.~Seager, J.~Jenkins) and the many TESS team members for their
efforts to make the mission a continued success.
%
%
%This study was based in part on observations at Cerro Tololo
%Inter-American Observatory at NSF's NOIRLab (NOIRLab Prop{.} ID
%2020A-0146; 2020B-0029 PI: Bouma), which is managed by the
%Association of Universities for Research in Astronomy (AURA) under a
%cooperative agreement with the National Science Foundation.
%
%
Finally, this research has made use of the Keck Observatory Archive (KOA),
which is operated by the W. M. Keck Observatory and the NASA Exoplanet
Science Institute (NExScI), under contract with the National
Aeronautics and Space Administration.  We also thank the Keck
Observatory staff for their support of HIRES and remote observing.  We
recognize the importance that the summit of Maunakea has within the
indigenous Hawaiian community, and are deeply grateful to have the
opportunity to conduct observations from this mountain.
%
% The Digitized Sky Survey was produced at the Space Telescope Science
% Institute under U.S. Government grant NAG W-2166.
% Figure~\ref{fig:scene} is based on photographic data obtained using
% the Oschin Schmidt Telescope on Palomar Mountain.
%

@@

\begin{figure*}[t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.7\textwidth]{f13.pdf}
	\end{center}
	\vspace{-0.7cm}
	\caption{
		{\bf Statistics of inter-flare arrival times}.  
    24 flares were recorded with amplitudes exceeding 0.5\% over the
    97.7 days of short cadence observations.  The histogram of the
    time intervals between every possible pair of flares is shown in
    black.  Some plausibly important timescales for star-planet
    interactions, namely the planetary orbital period and synodic
    period (the orbital period as seen from the rotating stellar frame) are
    shown along with their linear combinations.  Monte Carlo
    draws from a Poisson distribution are shown
    with the gray bands.  While peaks in the observed distribution do
    coincide with some of the ``special periods'', the
    statistical evidence for a non-Poissonian process driving the
    flares does not reach the $5$-$\sigma$ threshold.
		\label{fig:flarestats}
	}
\end{figure*}
